---
title: "Class10"
author: "Yuting Shen"
format: gfm
---

## 1. Importing candy data

```{r}
candy <-  read.csv("candy-data.txt", row.names=1)
head(candy)
```

> Q1. How many different candy types are in this dataset?

```{r}
dim(candy)
```

85 different candy types. 

> Q2. How many fruity candy types are in the dataset?

```{r}
sum(candy$fruity)
```
38 fruity candy types in the dataset. 

## 2. What is your favorate candy?

One of the most interesting variables in the dataset is `winpercent`. For a given candy this value is the percentage of people who prefer this candy over another randomly chosen candy from the dataset (what 538 term a matchup).

> Q3. What is your favorite candy in the dataset and what is it’s winpercent value?

My favorite candy is Twix. The winpercent is 81.6%. 

```{r}
rownames(candy)
```


```{r}
candy["Twix", ]$winpercent
```


> Q4. What is the winpercent value for “Kit Kat”?

```{r}
candy["Kit Kat", ]$winpercent
```

> Q5. What is the winpercent value for “Tootsie Roll Snack Bars”?


```{r}
candy["Tootsie Roll Snack Bars", ]$winpercent
```


There is a useful `skim()` function in the skimr package that can help give you a quick overview of a given dataset.


```{r}
skimr::skim(candy)
```

> Q6. Is there any variable/column that looks to be on a different scale to the majority of the other columns in the dataset?

Yes, the winpercent. 

> Q7. What do you think a zero and one represent for the candy$chocolate column?

True or False. zero means not chocolate and one means chocolate. 

> Q8. Plot a histogram of winpercent values

```{r}
hist(candy$winpercent)
```

```{r}
library(ggplot2)

ggplot(candy)+
  aes(winpercent)+
  geom_histogram(bins=10, col="red", fill="orange")
```

> Q9. Is the distribution of winpercent values symmetrical?

No 

> Q10. Is the center of the distribution above or below 50%?

The center of the distribution is below 50%. 

> Q11. On average is chocolate candy higher or lower ranked than fruit candy?

```{r}
choclate.inds <- as.logical(candy$chocolate)
choclate.wins <- candy[choclate.inds, ]$winpercent
mean(choclate.wins)
```
```{r}
fruity.inds <- as.logical(candy$fruity)
fruity.wins <- candy[fruity.inds, ]$winpercent
mean(fruity.wins)
```

On average, choclate is higher. 

Q12. Is this difference statistically significant?

The p-value is 2.871e-08, which is small. There is a significant difference. 
```{r}
t.test(choclate.wins, fruity.wins)

```

## 3. Overall Candy Rankings

> Q13. What are the five least liked candy types in this set?


```{r}
head(candy[order(candy$winpercent),], n=5)
```

```{r}
library("dplyr")
candy %>% arrange(winpercent) %>% head(5)
```


> Q14. What are the top 5 all time favorite candy types out of this set?

```{r}
candy %>% arrange(winpercent) %>% tail(5)
```

> Q15. Make a first barplot of candy ranking based on winpercent values.


```{r}
library(ggplot2)

ggplot(candy) + 
  aes(winpercent, rownames(candy)) +
  geom_col()
```

> Q16. This is quite ugly, use the reorder() function to get the bars sorted by winpercent?


```{r}
ggplot(candy) + 
  aes(winpercent, reorder(rownames(candy),winpercent)) +
  geom_col()
```

# set up some colors for diffrent candy types

```{r}
my_cols=rep("black", nrow(candy))
my_cols[as.logical(candy$chocolate)] = "chocolate"
my_cols[as.logical(candy$bar)] = "brown"
my_cols[as.logical(candy$fruity)] = "pink"
my_cols
```

```{r}
ggplot(candy) + 
  aes(winpercent, reorder(rownames(candy),winpercent)) +
  geom_col(fill=my_cols) 
```

```{r}
ggsave("tmp.png")
```

> Q17. What is the worst ranked chocolate candy?

Sixlets

> Q18. What is the best ranked fruity candy?

Starbust


## 4. Taking a look at pricepercent


What is the best (most  loked in terms of `winpercent`) for the money (in teerms of `pricepercent`)? 

To answer this I will make a plot of winpercent vs pricepercent

```{r}
library(ggrepel)

# How about a plot of price vs win
ggplot(candy) +
  aes(winpercent, pricepercent, label=rownames(candy)) +
  geom_point(col=my_cols) + 
  geom_text_repel(col=my_cols, size=2, max.overlaps = 5)
```

> Q19. Which candy type is the highest ranked in terms of winpercent for the least money - i.e. offers the most bang for your buck?

ReeseÕs Miniatures

> Q20. What are the top 5 most expensive candy types in the dataset and of these which is the least popular?

```{r}
ord <- order(candy$pricepercent, decreasing = TRUE)
head( candy[ord,c(11,12)], n=5 )
```

> Q21. Make a barplot again with geom_col() this time using pricepercent and then improve this step by step, first ordering the x-axis by value and finally making a so called “dot chat” or “lollipop” chart by swapping geom_col() for geom_point() + geom_segment().


```{r}
ggplot(candy) +
  aes(pricepercent, reorder(rownames(candy), pricepercent)) +
  geom_segment(aes(yend = reorder(rownames(candy), pricepercent), 
                   xend = 0), col="gray40") +
    geom_col()
```

```{r}
# Make a lollipop chart of pricepercent
ggplot(candy) +
  aes(pricepercent, reorder(rownames(candy), pricepercent)) +
  geom_segment(aes(yend = reorder(rownames(candy), pricepercent), 
                   xend = 0), col="gray40") +
    geom_point()
```


## 5 Exploring the correlation structure


```{r}
library(corrplot)
```

```{r}
cij <- cor(candy)
corrplot(cij)
```

> Q22. Examining this plot what two variables are anti-correlated (i.e. have minus values)?

Chocolate and fruity are anti-correlated. 

> Q23. Similarly, what two variables are most positively correlated?

Chocolate and winpercent are most positively correlated. 

## 6. Principal Component Analysis

```{r}
pca <- prcomp(candy, scale=TRUE)
summary(pca)
```

```{r}
plot(pca$x[,1:2], col=my_cols, pch=16)
```

```{r}
my_data <- cbind(candy, pca$x[,1:3])
```

```{r}
p <- ggplot(my_data) + 
        aes(x=PC1, y=PC2, 
            size=winpercent/100,  
            text=rownames(my_data),
            label=rownames(my_data)) +
        geom_point(col=my_cols)
p
```

```{r}
library(ggrepel)

p + geom_text_repel(size=3, col=my_cols, max.overlaps = 7)  + 
  theme(legend.position = "none") +
  labs(title="Halloween Candy PCA Space",
       subtitle="Colored by type: chocolate bar (dark brown), chocolate other (light brown), fruity (red), other (black)",
       caption="Data from 538")
```

```{r}
library(plotly)
```

```{r}
par(mar=c(8,4,2,2))
barplot(pca$rotation[,1], las=2, ylab="PC1 Contribution")
```

> Q24. What original variables are picked up strongly by PC1 in the positive direction? Do these make sense to you?

Fruity, hard and pluribus. Yes, it make sense. The candies belongs to these these three category are pushed to positive x-axis of the PC1. The result is consistent with the above corrplot. 